Theory about Normal Basis on $F(2^k)$

### 2017-03-03 dissertation/submission/unsat.tex
after submission: revision 2

The Boolean satisfiability (SAT) problem is the basis of 
most decidable decision problems, as well as the basis of many formal verification techniques. 
In this chapter, we discuss a special topic branching out from SAT theory. It is about a situation 
when SAT problems give negative answer, which are called unsatisfiability (UNSAT) problems. Within 
a set of constrains ({\it e.g.} clauses, formulas or polynomials) which is unsatisfiable, sometimes 
it is worthy to explore the reasons for UNSAT. From the execution of the GB algorithm, 
an auxiliary structure can be obtained to help explore the reasons causing UNSAT. This chapter introduces the 
details about the motivation, mechanism and implementation.

@@

\subsection{Preliminaries of SAT/UNSAT Theory}
The Boolean satisfiability (SAT) problem is a fundamental problem in computer science.
In the following part we define the terminology related to SAT problem \cite{anaICCAD}.

\begin{Definition}
A {\bf literal} $l$ is defined as a variable $v$, or its negation $\overline{v}$. A disjunction (OR relation)
of literals forms a {\bf clause}, {\it i.e}. $c = l_1\lor l_2 \lor \cdots \lor l_k$. 
A Boolean formula can always be written as {\bf conjunctive normal form} (CNF),
which is the conjunction (AND relation) of clauses: $F = c_1\land c_2\land\cdots\land c_k$.
\end{Definition}

@@

Figure \ref{fig:SAT} shows a simple example of SAT problem on circuit verification:
we need to verify whether subcircuit $A$ and subcircuit $B$ have the same function, so
we build a miter circuit for their outputs $X$ and $Y$, and the equivalence checking 
problem is turned into a SAT problem as follows:

@@

\subsection{Previous Work on CNF-UNSAT}
In most cases, SAT problems are modeled as conjunctive normal form (CNF) formulas, and solved by procedures based on 
Davis-Putnam and Davis-Logemann-Loveland (DPLL) algorithm. The main idea of DPLL algorithm
is recursive branching and backtrack searching. To improve the efficiency of DPLL-based
SAT solver, efforts are made on minimizing number of branches, accelerating unit propagation
and modifying backtrack algorithm. Recently state-of-art SAT solvers developed conflict-driven
clause learning (CDCL) technique to prune the search space, which is effective to reduce
search time.

When a SAT solver fails to give a SAT assignment, it will provide a {\it UNSAT proof} or {\it refutation proof}
to prove the problem is UNSAT. By analyzing clauses involved in the UNSAT proof, we can 
generate a subformula which remains UNSAT. A n\"aive method is to collect all leaf 
clauses in an UNSAT proof as an UNSAT core. In practice, a minimal UNSAT core is more valuable.
There are mainly two kinds of methods to find minimal UNSAT cores.
One is insertion-based method, which is achieved by adding clauses to the smallest subset 
until the subset turns to be UNSAT. The other method is deletion-based method,
which is realized by deleting clauses from a larger subset until the subset turns to be SAT.
Recently heuristics such as clause-set refinement \cite{modelrotation} and model rotation \cite{belov2011accelerating}
are applied as improvements on the deletion/insertion based methods.
These methods are also expanded to satisfiability modulo theories (SMT) \cite{cimatti2007simple}.

@@

The inspiration to use PC to solve SAT problems first came from \cite{ceiSTOC96}.
Using PC and its variations, researchers developed many SAT solvers \cite{STABLE,BLUEVERI,PolyBoRi}. 
Besides, researchers borrow concepts from PC, combine them with traditional DPLL and clause learning techniques
to build hybrid SAT solvers \cite{condratTACAS07,Zengler2010}.

\subsection{Exploiting UNSAT Cores for Abstraction Refinement}
The problem of finding small UNSAT cores has attracted interest for decades because of 
its applications to various of verification and synthesis problems.
In solving MaxSAT problems, a small UNSAT proof provides lower bound for the branch-and-bound searching algorithm \cite{li2009maxsat}.
It is applied to solving logic synthesis problems, such as Boolean function decomposition \cite{lee2008bi}. 
Small explanation generation in general constraint programming problems also rely on small UNSAT cores \cite{cambazard2008reformulating}.

UNSAT cores can find wide range of applications in circuit verification as well. Many abstraction refinement techniques require 
information mining from UNSAT proofs of intermediate abstractions. 
Here we use an abstraction refinement algorithm from \cite{zhang2005design} to explain how an UNSAT proof is
utilized in such techniques.

Bounded model checking (BMC) is a model checking technique which set upper bound to the length of all paths. 
It can be solved by solving a SAT problem. Given a model $M$, property $p$ and bound $k$, 
The $k$-BMC unrolls $M$ by $k$ clock-cycles and generates Boolean formula $F$ including violation check of $p$.
Then $F$ is fed to a SAT solver, if the SAT solver returns SAT, then $p$ is violated in 
some paths shorter than $k$. Otherwise, UNSAT indicates $p$ is not violated in all paths 
shorter than $k$. Algorithm \ref{alg:absrefine} makes an improvement on BMC using 
abstraction refinement.

@@

The first ``if-else" branch in Algorithm \ref{alg:absrefine} can be
explained as: we check if the conjunction of formulas 
$$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
is SAT or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
SAT, then a counterexample is found that violates property $p$. If the result is UNSAT, we cannot
assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
without any fix-point detection.
In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is UNSAT.
If there are some latch variables ($L_{abs}$) not included in this UNSAT core, then we can assert that the evaluations of 
these variables will not affect the unsatisfiability of original formula. Therefore, we can ignore them in the abstracted model.
In practice, we turn these latches into primary inputs/outputs as shown in Figure \ref{fig:absrefine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).

The second ``if-else" branch means: if we model check the abstracted machine $M'$ and find no
error trace, we can assert that property $p$ also holds on the original machine $M$. The reason of this assertion is that
the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an {\bf over-approximation}
of $M$, such that 
$$(M'\implies p) \implies (M\implies p)$$

If we find a violation on abstracted machine, then this
abstracted model is not a suitable model to check $p$, so we have to increase the bound $k$ to find a finer abstraction.

It is clear that UNSAT cores play an important role in abstraction refinement approaches. In \cite{zhang2005design}
the UNSAT core is extracted using a conventional CNF-SAT solver, which encounters the ``bit-blasting" problem
when the size of datapath (number of latches in Figure \ref{fig:absrefine}) is very large. \textbf{Here we propose an altogether new 
method based on Gr\"obner basis computation to extract UNSAT cores, and we believe it may become an efficient 
method according to the following observation:}

% {\it Gr\"obner basis is more suitable for UNSAT problems because of following theorem:}
{\it While the complexity of computing a GB over finite fields is exponential in the number of variables,
the GB  computation is observed to be more efficient for UNSAT problems.} 
The reason is as follows:
\begin{Theorem}[Weak Nullstellensatz]
Given ideal $J\subseteq \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure $\overline{\F}$
of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1".
$${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow \text{reduced GB}(J) = \{1\}$$
\end{Theorem}
It is well known that using Buchberger's algorithm and its variations to compute a GB has a very high space and
time complexity and is usually not practical. One reason is that the size of the GB may explode even if the term
ordering is carefully chosen. However if the reduced GB is $\{1\}$, which means every term in the original polynomials
will be canceled, the degree of remainders when computing GB with Buchberger's algorithm will be limited. 
Thus the number of polynomials in non-reduced GB is much smaller than usual.
Instead of applying polynomial calculus to SAT solving, it may be more efficient to try it for UNSAT
problems.

Moreover, conventional techniques are limited to bit-level variables (literals). Algebraic geometry methods
allow the use of word-level variables, which provides a strong potential for all applications which can be
modeled as polynomials in finite fields extensions.

@@

%% Let us first consider the problem of finding any $F_c \subset F$,
%% disregarding minimality. 
It is not hard to motivate that an UNSAT core should be identifiable
using the Gr\"obner basis algorithm: Assume that $F_c=F-\{f_j\}$. If
$GB(F) = GB(F_c) = \{1\}$, then it implies that $f_j$ is a member of
the ideal generated by $(F - \{f_j\})$, {\it i.e.} $f_j \in \langle F -
\{f_j\}\rangle$. Thus $f_j$ can be composed of the other
polynomials of $F_c$, so $f_j$ is  not a part of the UNSAT core, and
it can be safely discarded from $F_c$. This can be identified by means
of the GB algorithm for this ideal membership test.% (Definition \ref{def:gb}(ii)).

A n\"aive way (and inefficient way) to identify {\it a minimal core}
using the GB computation is as follows:
select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
({\it i.e.} if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
from the core; otherwise retain $f_i$ in $F_c$. Select a different
$f_i$ and continue until all polynomials $f_i$ are visited for
inclusion in $F_c$. This approach will produce a minimal core, as we
would have tested each polynomial $f_i$ for inclusion in the
core. This requires $O(|F|)$ calls to the GB engine, which is really
impractical.

\subsection{An Example Motivating our Approach}

Buchberger's algorithm picks pairs of polynomials from a given set, computes their $Spoly$, then reduces this $Spoly$
with the given set of polynomials. If the remainder is non-zero,
it is added to the set of polynomials.
By tracking $Spoly$ computations and multivariate divisions that lead to remainder 
1, we can obtain an UNSAT core. Moreover, we can identify a minimal UNSAT core with one-time
 execution of Buchberger's algorithm.

@@

\vspace{0.1in}
We compute its GB using Buchberger's algorithm with lexicographic term ordering $a>b>x>y>z$.
Since this problem is UNSAT, we will stop when ``1" is added to GB.
\begin{enumerate}[{1)}]
\item First we compute $Spoly(f_1,f_2)\xrightarrow{F}_{+} r_1$, remainder $r_1$ equals to $a$;
\item Update $F=F\cup r_1$;
\item Next we compute $Spoly(f_1,f_3)\xrightarrow{F}_{+} r_2$, remainder $r_2$ equals to $b$;
\item Update $F=F\cup r_2$;
\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Figure \ref{fig:originUNSAT}(a) shows;
\item Then we compute $Spoly(f_1,f_4) = s_3= a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
$r_1$ , the intermediate remainder $r_3 = b+1$. It can be immediately divided by $r_2$, and the remainder is ``1", we
terminate the Buchberger's algorithm;
\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Figure \ref{fig:originUNSAT}(b). 
From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. They constitute an UNSAT core for this problem
as these polynomials are the ``causes" of unsatisfiability of original set of polynomials.
\end{enumerate}
\end{Example}

\begin{figure}[bp]
\centering{
\includegraphics[width=\textwidth]{newfig/UNSAT.eps}}
\caption{DAG representing $Spoly$ computations and multivariate divisions.}
\label{fig:originUNSAT}
\end{figure}

@@

We investigate if it is possible to identify a core by analyzing the
$Spoly(f_i,f_j)\xrightarrow{F}_+ g_{ij}$ reductions in Buchberger's
algorithm. Since $F$ is itself an UNSAT core, 
definitely {\it there exists  
a sequence of $Spoly$ reductions in Buchberger's algorithm where
$Spoly(f_i, f_j) \xrightarrow{F}_+ 1$ is achieved.} Moreover, 
polynomial reduction algorithms can be suitably modified to record
which polynomials from $F$ are used in the division leading to
$Spoly(f_i,f_j)\xrightarrow{F}_+1$. This suggests that 
we should be able to identify a core by recording
the {\it data} generated by Buchberger's algorithm --- namely, the
critical pairs($f_i,f_j$) used in the $Spoly$ computations,
and the polynomials from $F$ used to cancel terms in the reduction
$Spoly(f_i,f_j)\xrightarrow{F}_+1$. The following example motivates
our approach to identify  $F_c \subseteq F$ using this data:

@@

Buchberger's algorithm terminates to a unique reduced GB, irrespective
of the order in which the critical pairs $(f_i,f_j)$ are selected and reduced by operation
$Spoly(f_i,f_j)\xrightarrow{F}_+g_{ij}$. Let us suppose that in the GB
computation corresponding to Example \ref{ex:1}, the first 3 critical
{\it Spoly} pairs analyzed are $(f_1, f_2), (f_3, f_4)$ and
$(f_2,f_5)$. It turns out that the $Spoly$-reductions corresponding to
these 3 pairs lead to the unit ideal. Recording the data
corresponding to this sequence of reductions is depicted by means of a
graph in Figure \ref{fig:refute}. We call this graph a {\it refutation tree}.

@@

In the figure, the nodes of the graph correspond to the polynomials
utilized in Buchberger's algorithm. The leaf nodes always correspond
to polynomials from the given generating set. An edge $e_{ij}$ from
node $i$ to node $j$ signifies that the polynomial at node $j$ results
from polynomial at node $i$. For example, consider the computation
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$, where $f_{10} = a + c +
1$. Since $Spoly(f_1, f_2) = f_1 - ac\cdot f_2$, the leaves
corresponding to $f_1$ and $ac\cdot f_2$ are created. The reduction
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$ is carried out as the
following sequence of 1-step divisions:
$Spoly(f_1,f_2)\xrightarrow{a\cdot f_2} ~\xrightarrow{f_3}~
\xrightarrow{c\cdot f_2}  ~\xrightarrow{f_2} f_{10}$. This is depicted
as the bottom subtree in the figure, terminating at polynomial
$f_{10}$. Moreover, the multiplication $a\cdot f_2$ implies that
division by $f_2$ resulted in the quotient $a$. The refutation tree of
Figure  \ref{fig:refute} shows further that
$Spoly(f_3,f_4)\xrightarrow{f_{10}} f_{11} = c+1$ and, finally,
$Spoly(f_5,f_2)\xrightarrow{f_{11}} 1$.

To identify an $F_c \subset F$, we start from the refutation node
``1'', and traverse the graph in reverse, all the way up to the
leaves. Then, all the leaves in the transitive fanin of ``1''
constitute an UNSAT core. The polynomials (nodes) that do not lie in
the transitive fanin of ``1'' can be safely discarded from $F_c$. From
Figure \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
as an UNSAT core of $F$.

@@

We record the S-polynomial reduction
$Spoly(f_i,f_j)\xrightarrow{F}_+{g_{ij}}$ that give a non-zero
remainder when divided by the system of polynomials $F$ at that
moment. The remainder $g_{ij}$ is a polynomial combination of
$Spoly(f_i,f_j)$ and the current basis $F$; thus, it can be
represented as:
\begin{equation}
\label{eqn1}
g_{ij}= S(f_i,f_j)+\displaystyle\sum_{k=1}^m c_kf_k,
\end{equation}
where $0\neq c_k\in\mathbb{F}[x_1,\ldots,x_d]$ and
$\{f_1,\ldots,f_m\}$ is the ``current'' system of polynomials
$F$. For each non-zero $g_{ij}$, we will record the following data:

\begin{equation}
\label{data1}
((g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{equation}

@@

Corresponding to Example \ref{ex:1} and the refutation tree shown in
Figure \ref{fig:refute}, we discover that the polynomial $f_3$ is used
only twice in the division process. In both occasions, the quotient of
the division is 1. From Figure \ref{fig:refute}, it follows that:
\begin{equation}
1 = (f_2 + f_5) + \dots + \mathbf{1\cdot f_3} + \dots + \mathbf{1\cdot
  f_3}+ \dots + (f_1 + f_2)
\end{equation}

@@

{\bf S-polynomial reduction:} We start computing the S-polynomials of the system of
generators $\{f_1,\ldots,f_s\}$, then divide each of them by the
current basis $G=\{f_1,\dots,f_s,\dots,f_m\}$, which is the
intermediate result of Buchberger's algorithm.  
In this way, we obtain expressions of the following type:
\begin{equation}
\label{eqn:red}
g_{ij}= \underbrace{h_{ij}f_{i}+h_{ji}f_{j}}_{Spoly(f_i,f_j)}+\displaystyle\sum_{k=1}^m c_kf_k
\end{equation}
If the remainder $g_{ij}$ is non-zero, we denote it by
$f_{m+1}$ and add it to the current set of generators $G$. We
also record the data as in Equation \ref{data1}: 
\begin{displaymath}
((f_{m+1}=g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{displaymath}
This data forms a part of the refutation tree rooted at node $f_{m+1}$.

>>>>

The Boolean satisfiability (SAT) problem is the basis of 
most decidable decision problems, as well as the basis of many formal verification techniques. 
In this chapter, we discuss a special topic branching out from SAT theory. It is about a situation 
when SAT problems give negative answer, which are called unsatisfiability (UNSAT) problems. Within 
a set of constrains (e.g., clauses, formulas, or polynomials) which is unsatisfiable, sometimes 
it is worthwhile to explore the reasons for UNSAT. From the execution of the GB algorithm, 
an auxiliary structure can be obtained to help explore the reasons causing UNSAT. This chapter introduces the 
details about the motivation, mechanism, and implementation.

@@

\subsection{Preliminaries of SAT/UNSAT Theory}
The Boolean satisfiability (SAT) problem is a fundamental problem in computer science.
In the following part we define the terminology related to a SAT problem \cite{anaICCAD}.

\begin{Definition}
A {\bf literal} $l$ is defined as a variable $v$, or its negation $\overline{v}$. A disjunction (OR relation)
of literals forms a {\bf clause}, i.e., $c = l_1\lor l_2 \lor \cdots \lor l_k$. 
A Boolean formula can always be written as {\bf conjunctive normal form} (CNF),
which is the conjunction (AND relation) of clauses: $F = c_1\land c_2\land\cdots\land c_k$.
\end{Definition}

@@

Figure \ref{fig:SAT} shows a simple example of a SAT problem on circuit verification:
we need to verify whether subcircuit $A$ and subcircuit $B$ have the same function, so
we build a miter circuit for their outputs $X$ and $Y$, and the equivalence checking 
problem is turned into a SAT problem as follows:

@@

\subsection{Previous Work on CNF-UNSAT}
In most cases, SAT problems are modeled as conjunctive normal form (CNF) formulas, and solved by procedures based on 
the Davis-Putnam and Davis-Logemann-Loveland (DPLL) algorithm. The main idea of the DPLL algorithm
is recursive branching and backtrack searching. To improve the efficiency of a DPLL-based
SAT solver, efforts are made to minimize the number of branches, accelerating unit propagation
and modifying backtrack algorithm. Recently, state-of-art SAT solvers developed a conflict-driven
clause learning (CDCL) technique to prune the search space, which is effective at reducing
search time.

When a SAT solver fails to give a SAT assignment, it will provide an {\it UNSAT proof} or {\it refutation proof}
to prove the problem is UNSAT. By analyzing clauses involved in the UNSAT proof, we can 
generate a subformula which remains UNSAT. A naive method is to collect all leaf 
clauses in an UNSAT proof as an UNSAT core. In practice, a minimal UNSAT core is more valuable.
There are mainly two kinds of methods to find minimal UNSAT cores.
One is the insertion-based method, which is achieved by adding clauses to the smallest subset 
until the subset turns to be UNSAT. The other method is the deletion-based method,
which is realized by deleting clauses from a larger subset until the subset turns to be SAT.
Recently heuristics such as clause-set refinement \cite{modelrotation} and model rotation \cite{belov2011accelerating}
have been applied as improvements on the deletion/insertion-based methods.
These methods are also expanded to satisfiability modulo theories (SMT) \cite{cimatti2007simple}.

@@

The inspiration to use PC to solve SAT problems first came from \cite{ceiSTOC96}.
Using PC and its variations, researchers developed many SAT solvers \cite{STABLE,BLUEVERI,PolyBoRi}. 
Besides, researchers borrow concepts from PC and combine them with traditional DPLL and clause learning techniques
to build hybrid SAT solvers \cite{condratTACAS07,Zengler2010}.

\subsection{Exploiting UNSAT Cores for Abstraction Refinement}
The problem of finding small UNSAT cores has attracted interest for decades because of 
its applications to various verification and synthesis problems.
In solving MaxSAT problems, a small UNSAT proof provides a lower bound for the branch-and-bound searching algorithm \cite{li2009maxsat}.
It is applied to solving logic synthesis problems, such as Boolean function decomposition \cite{lee2008bi}. 
Small explanation generation in general constraint programming problems also relies on small UNSAT cores \cite{cambazard2008reformulating}.

UNSAT cores can find a wide range of applications in circuit verification as well. Many abstraction refinement techniques require 
information mining from UNSAT proofs of intermediate abstractions. 
Here we use an abstraction refinement algorithm from \cite{zhang2005design} to explain how an UNSAT proof is
utilized in such techniques.

Bounded model checking (BMC) is a model checking technique which set an upper bound to the length of all paths. 
It can be solved by solving a SAT problem. Given a model $M$, property $p$, and bound $k$, 
the $k$-BMC unrolls $M$ by $k$ clock-cycles and generates Boolean formula $F$ including violation check of $p$.
Then $F$ is fed to a SAT solver, if the SAT solver returns SAT, then $p$ is violated in 
some paths shorter than $k$. Otherwise, UNSAT indicates $p$ is not violated in all paths 
shorter than $k$. Algorithm \ref{alg:absrefine} makes an improvement on BMC using 
abstraction refinement.

@@

The first ``if-else" branch in Algorithm \ref{alg:absrefine} can be
explained as: we check if the conjunction of formulas 
$$I(s_0)\land \bigwedge_{i=0}^{k-1}T(s_i,s_{i+1}) \land \neg p$$
is SAT or not, where $s_i$ denotes the set of reached states in $i$-th time-frame. If the result is
SAT, then a counterexample is found that violates property $p$. If the result is UNSAT, we cannot
assert that $p$ is satisfied for the original machine $M$ because we only unrolled $M$ for a given specific number of time-frames
without any fix-point detection.
In this algorithm, we analyze the UNSAT core composed by a set of clauses whose conjunction is UNSAT.
If there are some latch variables ($L_{abs}$) not included in this UNSAT core, then we can assert that the evaluations of 
these variables will not affect the unsatisfiability of the original formula. Therefore, we can ignore them in the abstracted model.
In practice, we turn these latches into primary inputs/outputs as shown in Figure \ref{fig:absrefine}(b) ($L_{abs} = \{l_1,\dots,l_m\}$).

The second ``if-else" branch means: if we model check the abstracted machine $M'$ and find no
error trace, we can assert that property $p$ also holds on the original machine $M$. The reason to this assertion is that
the abstracted states represented using abstracted latches cover the original states, which means $M'$ is an {\bf overapproximation}
of $M$, such that 
$$(M'\implies p) \implies (M\implies p)$$

If we find a violation on an abstracted machine, then this
abstracted model is not a suitable model to check $p$, so we have to increase the bound $k$ to find a finer abstraction.

It is clear that UNSAT cores play an important role in abstraction refinement approaches. In \cite{zhang2005design}
the UNSAT core is extracted using a conventional CNF-SAT solver, which encounters the ``bit-blasting" problem
when the size of the datapath (number of latches in Figure \ref{fig:absrefine}) is very large. Here we propose an altogether new 
method based on Gr\"obner basis computation to extract UNSAT cores, and we believe it may become an efficient 
method according to the following observation:

% {\it Gr\"obner basis is more suitable for UNSAT problems because of following theorem:}
{\it While the complexity of computing a GB over finite fields is exponential in the number of variables,
the GB  computation is observed to be more efficient for UNSAT problems.} 
The reason is as follows:
\begin{Theorem}[Weak Nullstellensatz]
Given ideal $J\subseteq \mathbb F[x_1,\dots,x_n]$, its variety over algebraic closure $\overline{\F}$
of field $\mathbb F$ is empty if and only if its reduced Gr\"obner basis contains only one generator ``1."
$${\bf V}_{\overline{\mathbb F}}(J) = \emptyset \Longleftrightarrow \text{reduced GB}(J) = \{1\}$$
\end{Theorem}
It is well known that using Buchberger's algorithm and its variations to compute a GB has a very high space and
time complexity and is usually not practical. One reason is that the size of the GB may explode even if the term
ordering is carefully chosen. However if the reduced GB is $\{1\}$, which means every term in the original polynomials
will be canceled, the degree of remainders when computing GB with Buchberger's algorithm will be limited. 
Thus the number of polynomials in nonreduced GB is much smaller than usual.
Instead of applying polynomial calculus to SAT solving, it may be more efficient to try it for UNSAT
problems.

Moreover, conventional techniques are limited to bit-level variables (literals). Algebraic geometry methods
allow the use of word-level variables, which provides a strong potential for all applications which can be
modeled as polynomials in finite-field extensions.

@@

%% Let us first consider the problem of finding any $F_c \subset F$,
%% disregarding minimality. 
It is not hard to figure out that an UNSAT core should be identifiable
using the Gr\"obner basis algorithm: Assume that $F_c=F-\{f_j\}$. If
$GB(F) = GB(F_c) = \{1\}$, then it implies that $f_j$ is a member of
the ideal generated by $(F - \{f_j\})$, {\it i.e.} $f_j \in \langle F -
\{f_j\}\rangle$. Thus $f_j$ can be composed of the other
polynomials of $F_c$, so $f_j$ is  not a part of the UNSAT core, and
it can be safely discarded from $F_c$. This can be identified by means
of the GB algorithm for this ideal membership test.% (Definition \ref{def:gb}(ii)).

A naive way (and inefficient way) to identify {\it a minimal core}
using the GB computation is as follows:
Select a polynomial $f_i$ and see if $V(F_c - \{f_i\}) = \emptyset$
({\it i.e.} if reduced $GB(F_c - \{f_i\}) = \{1\}$). If so, discard $f_i$
from the core; otherwise retain $f_i$ in $F_c$. Select a different
$f_i$ and continue until all polynomials $f_i$ are visited for
inclusion in $F_c$. This approach will produce a minimal core, as we
would have tested each polynomial $f_i$ for inclusion in the
core. This requires $O(|F|)$ calls to the GB engine, which is really
impractical.

\subsection{An Example Motivating Our Approach}

Buchberger's algorithm picks pairs of polynomials from a given set, computes their $Spoly$, then reduces this $Spoly$
with the given set of polynomials. If the remainder is nonzero,
it is added to the set of polynomials.
By tracking $Spoly$ computations and multivariate divisions that lead to remainder 
1, we can obtain an UNSAT core. Moreover, we can identify a minimal UNSAT core with one-time
 execution of Buchberger's algorithm.

@@

\vspace{0.1in}
We compute its GB using Buchberger's algorithm with lexicographic term ordering $a>b>x>y>z$.
Since this problem is UNSAT, we will stop when ``1" is added to GB.
\begin{enumerate}[{1)}]
\item First we compute $Spoly(f_1,f_2)\xrightarrow{F}_{+} r_1$, remainder $r_1$ equals to $a$;
\item Update $F=F\cup r_1$;
\item Next we compute $Spoly(f_1,f_3)\xrightarrow{F}_{+} r_2$, remainder $r_2$ equals to $b$;
\item Update $F=F\cup r_2$;
\item We can use a directed acyclic graph (DAG) to represent the process to get $r_1,r_2$, as Figure \ref{fig:originUNSAT}(a) shows;
\item Then we compute $Spoly(f_1,f_4) = s_3= a+b+1$, obviously $a+b+1$ can be reduced (multivariate divided) by
$r_1$ , the intermediate remainder $r_3 = b+1$. It can be immediately divided by $r_2$, and the remainder is ``1," we
terminate the Buchberger's algorithm;
\item We draw a DAG depicting the process through which we obtain remainder ``1" as shown in Figure \ref{fig:originUNSAT}(b). 
From leaf ``1" we backtrace the graph to roots $f_1,f_2,f_3,f_4$. They constitute an UNSAT core for this problem
as these polynomials are the ``causes" of the unsatisfiability of the original set of polynomials.
\end{enumerate}
\end{Example}

\begin{figure}[tbp]
\centering{
\includegraphics[width=\textwidth]{newfig/UNSAT.eps}}
\caption{DAG representing $Spoly$ computations and multivariate divisions.}
\label{fig:originUNSAT}
\end{figure}

@@

We investigate whether it is possible to identify a core by analyzing the
$Spoly(f_i,f_j)\xrightarrow{F}_+ g_{ij}$ reductions in Buchberger's
algorithm. Since $F$ is itself an UNSAT core, 
definitely {\it there exists  
a sequence of $Spoly$ reductions in Buchberger's algorithm where
$Spoly(f_i, f_j) \xrightarrow{F}_+ 1$ is achieved.} Moreover, 
polynomial reduction algorithms can be suitably modified to record
which polynomials from $F$ are used in the division leading to
$Spoly(f_i,f_j)\xrightarrow{F}_+1$. This suggests that 
we should be able to identify a core by recording
the {\it data} generated by Buchberger's algorithm --- namely, the
critical pairs($f_i,f_j$) used in the $Spoly$ computations,
and the polynomials from $F$ used to cancel terms in the reduction
$Spoly(f_i,f_j)\xrightarrow{F}_+1$. The following example motivates
our approach to identify  $F_c \subseteq F$ using this data:

@@

Buchberger's algorithm terminates to a unique reduced GB, irrespective
of the order in which the critical pairs $(f_i,f_j)$ are selected and reduced by operation
$Spoly(f_i,f_j)\xrightarrow{F}_+g_{ij}$. Let us suppose that in the GB
computation corresponding to Example \ref{ex:1}, the first 3 critical
{\it Spoly} pairs analyzed are $(f_1, f_2), (f_3, f_4)$, and
$(f_2,f_5)$. It turns out that the $Spoly$-reductions corresponding to
these 3 pairs lead to the unit ideal. Recording the data
corresponding to this sequence of reductions is depicted by means of a
graph in Figure \ref{fig:refute}. We call this graph a {\it refutation tree}.

@@

In the figure, the nodes of the graph correspond to the polynomials
utilized in Buchberger's algorithm. The leaf nodes always correspond
to polynomials from the given generating set. An edge $e_{ij}$ from
node $i$ to node $j$ signifies that the polynomial at node $j$ results
from the polynomial at node $i$. For example, consider the computation
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$, where $f_{10} = a + c +
1$. Since $Spoly(f_1, f_2) = f_1 - ac\cdot f_2$, the leaves
corresponding to $f_1$ and $ac\cdot f_2$ are created. The reduction
$Spoly(f_1,f_2)\xrightarrow{F}_+ f_{10}$ is carried out as the
following sequence of 1-step divisions:
$Spoly(f_1,f_2)\xrightarrow{a\cdot f_2} ~\xrightarrow{f_3}~
\xrightarrow{c\cdot f_2}  ~\xrightarrow{f_2} f_{10}$. This is depicted
as the bottom subtree in the figure, terminating at polynomial
$f_{10}$. Moreover, the multiplication $a\cdot f_2$ implies that
division by $f_2$ resulted in the quotient $a$. The refutation tree of
Figure  \ref{fig:refute} shows further that
$Spoly(f_3,f_4)\xrightarrow{f_{10}} f_{11} = c+1$ and, finally,
$Spoly(f_5,f_2)\xrightarrow{f_{11}} 1$.

To identify an $F_c \subset F$, we start from the refutation node
``1'' and traverse the graph in reverse, all the way up to the
leaves. Then, all the leaves in the transitive fanin of ``1''
constitute an UNSAT core. The polynomials (nodes) that do not lie in
the transitive fanin of ``1'' can be safely discarded from $F_c$. From
Figure \ref{fig:refute}, $F_c = \{f_1,f_2,f_3,f_4,f_5\}$ is identified
as an UNSAT core of $F$.

@@

We record the S-polynomial reduction
$Spoly(f_i,f_j)\xrightarrow{F}_+{g_{ij}}$, which gives a nonzero
remainder when divided by the system of polynomials $F$ at that
moment. The remainder $g_{ij}$ is a polynomial combination of
$Spoly(f_i,f_j)$ and the current basis $F$; thus, it can be
represented as
\begin{equation}
\label{eqn1}
g_{ij}= S(f_i,f_j)+\displaystyle\sum_{k=1}^m c_kf_k,
\end{equation}
where $0\neq c_k\in\mathbb{F}[x_1,\ldots,x_d]$ and
$\{f_1,\ldots,f_m\}$ is the ``current'' system of polynomials
$F$. For each nonzero $g_{ij}$, we will record the following data: 
\begin{equation}
\label{data1}
((g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{equation}

@@

Corresponding to Example \ref{ex:1} and the refutation tree shown in
Figure \ref{fig:refute}, we discover that the polynomial $f_3$ is used
only twice in the division process. In both occasions, the quotient of
the division is 1. From Figure \ref{fig:refute}, it follows that
\begin{equation}
1 = (f_2 + f_5) + \dots + \mathbf{1\cdot f_3} + \dots + \mathbf{1\cdot
  f_3}+ \dots + (f_1 + f_2)
\end{equation}

@@

{\bf S-polynomial reduction:} We start computing the S-polynomials of the system of
generators $\{f_1,\ldots,f_s\}$, then divide each of them by the
current basis $G=\{f_1,\dots,f_s,\dots,f_m\}$, which is the
intermediate result of Buchberger's algorithm.  
In this way, we obtain expressions of the following type:
\begin{equation}
\label{eqn:red}
g_{ij}= \underbrace{h_{ij}f_{i}+h_{ji}f_{j}}_{Spoly(f_i,f_j)}+\displaystyle\sum_{k=1}^m c_kf_k
\end{equation}
If the remainder $g_{ij}$ is nonzero, we denote it by
$f_{m+1}$ and add it to the current set of generators $G$. We
also record the data as in Equation \ref{data1}: 
\begin{displaymath}
((f_{m+1}=g_{ij})(f_{i},h_{ij})(f_{j},h_{ji})| (c_{k1},f_{k1}),(c_{k2},f_{k2}),\dots,(c_{kl},f_{kl}))
\end{displaymath}
This data forms a part of the refutation tree rooted at node $f_{m+1}$.

{\bf Recording the coefficients:} In Equation \ref{eqn:red} we obtain a
vector of polynomial coefficients $c_k$ where $k>s$. These
coefficients are associated with new elements (remainders) in the
Gr\"obner basis that are not a part of the UNSAT core. 
%which cannot benefit the UNSAT core extraction. 
Since each polynomial $f_k$, ($k>s$) is generated by
$\{f_1,\dots,f_s\}$, we can re-express $f_k$ in terms of $\{f_1,\dots,
f_s\}$. Thus, each $f_k, k>s$ can be written as $f_k = d_1f_1 + \dots
+ d_sf_s$. This process adds a new row $(d_1,\dots,d_s)$ to the
coefficient matrix $M$.
